---
title: "Manuscript code for 'Small Arctic rivers transport legacy contaminants to coastal areas in Kongsfjorden, Svalbard'"
author: "Maeve McGovern"
date: "September, 2020"
output: html_document
---

This file contains the code used for the tables and figures in the manuscript 'Small Arctic rivers transport legacy contaminants to coastal areas in Kongsfjorden, Svalbard'

```{r setup, include=F}
knitr::opts_chunk$set(echo=T, message=F, warning=F)

# For data prep
library(tidyverse)

# For stats and plotting
library(vegan)
library(broom)
library(PlotSvalbard)

# For plots and table formatting
library(patchwork)
library(kableExtra)
library(RColorBrewer)
library(pdftools)
library(jpeg)
library(ggrepel)
set.seed(42)

# Load datafile
Data <- read_csv("Secondary_Data/KongsPOPs_Cleaned_Data.csv") %>% 
   select(-Date) %>% 
   mutate(Name = case_when(grepl("Bay", Station) ~ "BE",
                           grepl("Kva", Station) ~"KH", 
                           grepl("Gas", Station) ~"GA", 
                           grepl("Inner", Station) ~"IF", 
                           grepl("Kongs", Station) ~"KV"))
  
# Load secondary datafiles
LogKow <- read_csv("Secondary_Data/LogKOW.csv") # Literature values for logKOWs used in Figure 6. 
CTD <- read_csv('Secondary_Data/TS_diagram.csv') # CTD datafiles for the supplemental temperature-salinity diagram
```

## Figure 1. Station Map

```{r Figure 1}
# Station Map using Mikko's PlotSvalbard package
# Vihtakari, M. (2019). PlotSvalbard: PlotSvalbard – Plot Research Data From Svalbard on Maps. Rpackage version 0.8.5.

Kongsfjord <- Data %>% 
  select(Station, Name, Longitude, Latitude) %>% 
  distinct()

Kongsfjord$Name <- factor(Kongsfjord$Name, levels = c("KV","IF", "GA", "BA","KH"))

x <- transform_coord(Kongsfjord, lon = "Longitude", lat = "Latitude", bind = TRUE)

basemap("kongsfjorden", limits = c(11.3, 12.69, 78.86, 79.1), 
          bathymetry=TRUE, round.lat = 0.05, round.lon = 0.5) +
  geom_point(data = x, aes(x = lon.utm, y = lat.utm, color=Name), cex = 10) + 
  scale_color_brewer(palette="PuBuGn") +
  geom_text(data = x, aes(x = lon.utm + 5, y = lat.utm + 5, 
                          label = Name), cex= 5, fontface = 2) +
  theme(legend.position='none')
```


## Figure 2. Boxplot of concentrations for main groups. 

```{r Figure 2}

# Concentrations of contamiant groups by type (marine vs. river) and fraction (particulate vs. dissolved). 

Data$Type <- factor(Data$Type,levels = c("River", "Fjord"))

S1 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = Sum_PCB, fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + 
  xlab(NULL) + ylab("SUM PCB") + theme_classic() +
  scale_fill_manual(values=c("#8B9DAF","lightsalmon4"))  + 
  theme(legend.position='none') +
  geom_point(position = position_jitterdodge(), stat ='identity', shape = 21, size = 2, alpha = 1) + 
  scale_y_log10() + xlab(NULL) + ylab(expression(paste(Sigma, 'PCB' ~ (pg~L^-1)))) + 
  theme(text = element_text(size = 18))

S2 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = Sum_DDTs, fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + xlab(NULL) +
  scale_fill_manual(values = c("#8B9DAF","lightsalmon4")) + theme_classic() + 
  theme(legend.position ='none') +
  geom_point(position=position_jitterdodge(), stat='identity', shape = 21, size = 2,alpha=1) + 
  scale_y_log10() + ylab(expression(paste(Sigma, 'DDT' ~ (pg~L^-1)))) + xlab(NULL) + 
  theme(text = element_text(size = 18))

S3 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = Sum_HCH, fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + xlab(NULL) + ylab("SUM HCH") +
  scale_fill_manual(values = c("#8B9DAF","lightsalmon4")) + theme_classic() + 
  theme(legend.position='none') +
  geom_point(position = position_jitterdodge(), stat ='identity', shape = 21, size = 2, alpha =1) + 
  scale_y_log10() + ylab(expression(paste(Sigma, 'HCH' ~ (pg~L^-1)))) + 
  xlab(NULL) + theme(text = element_text(size = 18))

S4 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = Sum_Chlordanes,fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + xlab(NULL) + ylab("SUM Chlordane") + 
  scale_fill_manual(values = c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + theme(legend.position ='none') +
  geom_point(position = position_jitterdodge(), stat='identity', shape = 21, size = 2,alpha=1) + 
  scale_y_log10() + ylab(expression(paste(Sigma, 'Chlordanes' ~ (pg~L^-1)))) + 
  xlab(NULL) + theme(text = element_text(size=18))

S5 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = HCB,fill=Fraction)) + 
  geom_boxplot(outlier.shape = NA) + xlab(NULL) +
  scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + theme(legend.position='none') +
  geom_point(position = position_jitterdodge(), stat ='identity', shape = 21, size = 2, alpha = 1) + 
  scale_y_log10() + ylab(expression(paste('HCB' ~ (pg~L^-1)))) + 
  xlab(NULL) + theme(text = element_text(size = 18))

S6 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x=Type, y=PeCB, fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + xlab(NULL) +
  scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + theme(legend.position ='none') +
  geom_point(position = position_jitterdodge(), stat ='identity', shape = 21, size = 2, alpha = 1) + 
  scale_y_log10() + ylab(expression(paste('PeCB' ~ (pg~L^-1)))) + 
  xlab(NULL) + theme(text = element_text(size=18))

S7 <- Data %>%
  filter(!Volume_L == 1) %>%
  arrange(Lab_ID, Volume_L) %>%
  ggplot(aes(x = Type, y = Sum_OC, fill = Fraction)) + 
  geom_boxplot(outlier.shape = NA) + 
  xlab(NULL) + scale_fill_manual(values = c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + theme(legend.position ='none') +
  geom_point(position = position_jitterdodge(), stat ='identity', shape = 21, size = 2, alpha = 1) +
  ylab(expression(paste(Sigma, 'OCs' ~ (pg~L^-1)))) + xlab(NULL) + 
  theme(text = element_text(size=18)) + scale_y_log10()


Figure2 <- ((S7|S1|S3)/(S4|S5|S2) + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') & theme(legend.position = 'bottom'))


# Save plot
ggsave("Figure2.pdf", Figure2, width = 12, height = 8.8, device = cairo_pdf)
```


```{r Concentration Tests}

# Normality and wilcoxon rank-sum tests of groups of contaminants between river vs. fjord for dissolved and particulate fractions. 

# Normality tests. 
Shapiro <- Data %>% 
  select(Type, Fraction, HCB, Sum_PCB, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_OC) %>% 
  pivot_longer(cols = -c(Type, Fraction), names_to = "Compound", values_to= 'Concentration') %>%   
  nest(-Compound) %>% 
  mutate(shapiro = map(data, ~shapiro.test(.x$Concentration)), 
         glanced = map(shapiro, glance), tidied = map(shapiro, tidy)) %>% 
  unnest(glanced) %>% 
  select(Compound, W = statistic, p.value) 


# Table of non-parametric Wilcoxon rank-sum tests 
Dis <- Data %>%
  filter(Fraction=='Dissolved')  %>%
  select(c(Type, HCB, Sum_PCB, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_OC))  %>%
  pivot_longer(cols = -c(Type), names_to = "Compound", values_to = 'Concentration') %>%   
  nest(-Compound) %>%                  
  mutate(results = map(data, ~tidy(wilcox.test(Concentration ~ Type, data = .)))) %>%
  select(-c(data)) %>%
  unnest(results)

Part <- Data %>%
  filter(Fraction=='Particulate')  %>%
  select(c(Type,HCB, Sum_PCB, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_OC))  %>%
  pivot_longer(cols = -c(Type), names_to = "Compound", values_to= 'Concentration') %>%   
  nest(-Compound) %>%                  
  mutate(results = map(data, ~tidy(wilcox.test(Concentration ~ Type, data = .)))) %>%
  select(-c(data)) %>%
  unnest(results)

Dis %>% 
  select(-c(method, alternative))  %>% 
  kable(caption= 'Dissolved: Results of Wilcoxon Rank Sum Two-sided Test btw Fjord and River Dissolved samples', digits=4) %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE)

Part  %>% 
  select(-c(method, alternative))  %>% 
  kable(caption= 'Particulate: Results of Wilcoxon Rank Sum Two-sided Test btw Fjord and River Particulate samples', digits=4) %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE)


PCB7 <- Data %>% 
  ungroup() %>% 
  filter(Volume_L != "1") %>%
  group_by(Station, Type, Fraction, Volume_L) %>% 
  mutate(Sum_PCB7 = sum(c_across(c(PCB_28, PCB_52, PCB_101, PCB_118, PCB_138, PCB_153, PCB_180)))) %>% 
  select(c(Station, Type, Fraction, Sum_PCB7)) %>% 
  ungroup() %>% 
  group_by(Type, Fraction) %>% 
  summarise(n=n(), 
            Mean = mean(Sum_PCB7, na.rm=T), 
            stdev = sd(Sum_PCB7, na.rm=T), 
            min = min(Sum_PCB7, na.rm=T), 
            max = max(Sum_PCB7, na.rm=T))

PCB7_Table <- PCB7 %>% 
kable(caption= 'KongsPOPs PCB7 (PCB-28, PCB-52, PCB-101, PCB-118, PCB-138, PCB-153, PCB-180)', digits=2, booktabs = T, align = c("r"))  %>% 
kable_classic(full_width = F)
```


## Figure 3. Multivariate analyses
 
RDA (Concentrations) and CCA (Composition) constrained by type (Rivers vs. Fjord) for the dissolved fraction

```{r Figure 3}

# Figure 3

#Figure 3b: Concentration
# PCA/RDA 

RDA <- Data %>% 
  filter(Fraction=='Dissolved') %>% 
  filter(Volume_L !=1) %>% 
  select(-c(Station, Longitude, Latitude, Fraction, Volume_L, Lab_ID)) %>% 
  relocate(Name, Type)

Concentrations_rep <- RDA %>% select(c(aHCH:PCB_187)) 
Name <- RDA %>% select(Name)
Sums <- RDA %>% select(c(Sum_PCB, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_Chlorobenzenes, Sum_OC))
Colors <- brewer.pal(5,'PuBuGn')

#Dissolved Concentrations
Tot_log = log10(Concentrations_rep+1)
Tot_logs <- scale(Tot_log, center = TRUE, scale = FALSE)
rda.Tot <- rda(Tot_logs ~ Type, data = RDA)
rda.Tot
anova(rda.Tot, by='margin')

RDAscores <- scores(rda.Tot, display = "sites") %>% 
  as.data.frame() 
RDAvect <- scores(rda.Tot, display = "species") %>% 
  as.data.frame()

RDA$Name <- factor(RDA$Name, levels = c("KV", "IF", "GA", "BE","KH"))

# Envfits
fit <- envfit(rda.Tot, Sums, perm = 999, na.rm=T) %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species")

fit$species <- factor(fit$species, levels = c("Sum_OC","Sum_PCB", "Sum_HCH", "Sum_Chlordanes","Sum_DDTs","Sum_Chlorobenzenes"),
       ordered = TRUE,
       labels = c(expression(paste(Sigma,'OCs')),
            expression(paste(paste(Sigma, "PCB" [17]))), 
            expression(paste(Sigma,'HCHs')),
            expression(paste(Sigma,'Chlordanes')),
            expression(paste(Sigma,'DDTs')),
            expression(paste(Sigma,'Chlorobenzenes'))))

Plot1 <- ggplot() +
  geom_point(data = RDAscores, aes(x = RDA1, y = PC1, fill = RDA$Name, shape = RDA$Type), 
             color= "gray45", size= 6.5) +
  scale_fill_manual(values=c(Colors)) + 
  theme_classic() + 
  scale_shape_manual(values=c(23,21)) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) + 
  ylim(-1.2, 1.2) + xlim(-1.2,1.2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2)  +
  labs(x = "RDA1 (50 %)", y = "PC1 (37 %)") +  
  geom_segment(data = fit, aes(x = 0, xend = RDA1, y = 0, yend = PC1), 
               arrow = arrow(length = unit(0.25, "cm")), col = "black") +
  geom_text_repel(data = fit, aes(x = RDA1, y= PC1, label = species), 
                  parse = T,col="black", size=4.5) +
  theme(text = element_text(size=17), legend.text = element_text(size=rel(0.6)), 
        legend.title = element_blank(), panel.background = element_rect(colour = "black", size=1), 
        legend.position = "none") + 
  geom_text(data = RDAscores, aes(x = RDA1, y= PC1, label = RDA$Name), 
            size= 3, parse = T, col="black") +
  guides(fill = guide_legend(override.aes=list(shape=21)), 
         shape = guide_legend(override.aes=list(color="black"))) 


#Figure 4b. Composition
#Code for CA/CCA 

CCA <- Data %>% 
   filter(Fraction=='Dissolved') %>% 
   mutate_at(vars(c(aHCH, bHCH, gHCH,PCB_18, PCB_28, PCB_31, PCB_33, PCB_47, PCB_52, 
                   PCB_66, PCB_74, PCB_99, PCB_101, PCB_105, PCB_118, PCB_138, PCB_149, 
                   PCB_153, PCB_180, PCB_187,trans_Chlordane, cis_Chlordane, trans_Nonachlor, 
                   cis_Nonachlor,ppDDE, opDDD, ppDDD, opDDT, ppDDT, HCB, PeCB)), ~(./`Sum_OC`))  %>% 
   mutate(Sum_PCB = PCB_18 + PCB_28 + PCB_31 + PCB_33 + PCB_47 + PCB_52 + PCB_66 + 
         PCB_74 + PCB_99 + PCB_101 + PCB_105 + PCB_118 + PCB_138 + PCB_149 + 
           PCB_153 + PCB_180 + PCB_187) %>%
   mutate(Sum_PCB7=PCB_28+PCB_52+PCB_101+PCB_138+PCB_153+PCB_180) %>%
   mutate(Sum_HCH= aHCH+bHCH+gHCH) %>%
   mutate(Sum_DDTs = ppDDE+opDDD+ppDDD+opDDT+ppDDT) %>%
   mutate(Sum_Chlordanes = trans_Chlordane+cis_Chlordane+trans_Nonachlor+cis_Nonachlor) %>%
   mutate(Sum_Chlorobenzenes = HCB+PeCB) %>%
   mutate(Sum_OC = Sum_Chlordanes+Sum_HCH+Sum_DDTs+Sum_PCB+PeCB+HCB) %>%
   select(-c(Station, Longitude, Latitude, Fraction, Volume_L, Lab_ID, Sum_OC)) %>% 
   relocate(Name, Type)
  
Composition <- CCA %>% select(c(aHCH:PCB_187)) 
Name <- CCA %>% select(Name)
Sums <- CCA %>% select(c(Sum_PCB, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_Chlorobenzenes))
Colors <- brewer.pal(5,'PuBuGn')

CCA$Name <- factor(CCA$Name, levels = c("KV", "IF", "GA", "BE","KH"))

cca.Tot <- cca(Composition ~ Type, data=CCA)
cca.Tot
anova(cca.Tot, by='margin')

CCAscores <- scores(cca.Tot, display = "sites") %>% 
  as.data.frame() 
CCAvect <- scores(cca.Tot, display = "species") %>% 
  as.data.frame()

# Envfits
fit2 <- envfit(cca.Tot, Sums, perm = 999, na.rm=T) %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species")

fit2$species <- factor(fit2$species, levels = c("Sum_PCB", "Sum_HCH", "Sum_Chlordanes","Sum_DDTs","Sum_Chlorobenzenes"),
       ordered = TRUE,
       labels = c(expression(paste(paste(Sigma, "PCB" [17]))), 
            expression(paste(Sigma,'HCHs')),
            expression(paste(Sigma,'Chlordanes')),
            expression(paste(Sigma,'DDTs')),
            expression(paste(Sigma,'Chlorobenzenes'))))
#Plot
Plot2 <- ggplot() +
  geom_point(data = CCAscores, aes(x = CCA1, y = CA1, fill = CCA$Name, shape = CCA$Type), 
             color= "gray45",size= 6.5) +
  scale_fill_manual(values=c(Colors)) + 
  theme_classic() + 
  scale_shape_manual(values=c(23,21)) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) + 
  ylim(-2.5, 2.5) + xlim(-2.5, 2.5) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2)  +
  labs(x = "CCA1 (28 %)", y = "CA1 (43 %)") +  
  geom_segment(data = fit2, aes(x = 0, xend = CCA1, y = 0, yend = CA1), 
               arrow = arrow(length = unit(0.25, "cm")), col = "black") +
  geom_text_repel(data = fit2, aes(x = CCA1, y= CA1, label = species), 
                  parse = T,col="black", size=4.5) +
  theme(text = element_text(size = 17), legend.text = element_text(size=rel(0.6)), 
        legend.title = element_blank(), panel.background = element_rect(colour = "black", size=1),
        legend.position = "none") + 
  geom_text(data = CCAscores, aes(x = CCA1, y= CA1, label = CCA$Name), 
            size= 3, parse = T, col="black") +
  guides(fill = guide_legend(override.aes=list(shape=21)), 
         shape = guide_legend(override.aes=list(color="black")))

Figure3 <- Plot1/Plot2

# Save plot
ggsave("Figure3.pdf", Figure3, width = 5, height = 8.8, device = cairo_pdf)
```


## Figure 4. Stacked barplot of composition and concentrations for main groups. 

```{r Figure 4 Dataprep}
# Compositional barplots allow us to inlcude samples where sampling volume was missing. First we assign samples to 'replicates' of each station (rep1-rep4) and take the average. Then I use the average concentration to make the compositional barplot and lay the replicates on top as concentrations on the second axis. 

# Data preparation
Reps <- Data %>%
  select(-c(PCB_18, PCB_31, PCB_33, PCB_47, PCB_66, PCB_74, PCB_99, 
            PCB_105, PCB_118, PCB_149, PCB_187, Volume_L, Latitude, Longitude, Station)) %>%
  ungroup() %>%
  pivot_longer(cols=-c(Name, Type, Fraction, Lab_ID), names_to = "Compound", values_to= 'Concentration') %>%
  mutate(Reps= 'Rep1') %>%
  mutate(Reps= ifelse(grepl("3075|3076|3083|3084|3091|3092|3095|3096|3099|3100|3107|3108|3113|3114",
                            Lab_ID),'Rep2', Reps)) %>%
  mutate(Reps= ifelse(grepl("3077|3078|3085|3086|3109|3110", Lab_ID), 'Rep3', Reps)) %>%
  mutate(Reps= ifelse(grepl("3079|3080|3087|3088", Lab_ID), 'Rep4', Reps)) %>%
  select(-Lab_ID) %>% 
  pivot_wider(names_from = "Reps", values_from= 'Concentration') %>%
  mutate(Average = rowMeans(select(.,starts_with("R")), na.rm = TRUE))

Reps2 <- Data %>% 
  select(-c(PCB_18, PCB_31, PCB_33, PCB_47, PCB_66, PCB_74, PCB_99, PCB_105, PCB_118, 
            PCB_149, PCB_187, Latitude, Longitude, Station)) %>%
  filter(Volume_L!=1) %>% 
  arrange(Lab_ID, Volume_L) %>%
  ungroup() %>%
  pivot_longer(cols=-c(Lab_ID, Name, Type, Fraction, Volume_L), names_to = "Compound", values_to= 'Concentration') %>%
  mutate(Reps= 'Rep1') %>%
  mutate(Reps= ifelse(grepl("3075|3076|3083|3084|3091|3092|3095|3096|3099|3100|3107|3108|3113|3114",
                            Lab_ID),'Rep2', Reps)) %>%
  mutate(Reps= ifelse(grepl("3077|3078|3085|3086|3109|3110",
                            Lab_ID), 'Rep3', Reps)) %>%
  mutate(Reps= ifelse(grepl("3079|3080|3087|3088",
                            Lab_ID), 'Rep4', Reps)) %>%
  select(-c(Lab_ID, Volume_L)) %>%
  pivot_wider(names_from = "Reps", values_from= 'Concentration')  %>%
  filter(Compound!= 'Sum_OC') %>% 
  mutate(Group = case_when(grepl("PCB", Compound) ~ "PCB", 
                           grepl("DD", Compound) ~ "DDT", 
                           grepl("HCH", Compound) ~ "HCH", 
                           grepl("hl",  Compound) ~ "Chlordanes", 
                           grepl("Sum", Compound) ~ "Sums"))  %>%
  mutate(Group = ifelse(grepl("Sum",Compound), 'Sums', Group)) %>% 
  group_by(Name, Type, Fraction, Group) %>% 
  mutate(Rep1_c = sum(Rep1)) %>% 
  mutate(Rep2_c = sum(Rep2)) %>% 
  mutate(Rep3_c = sum(Rep3)) %>% 
  mutate(Rep4_c = sum(Rep4)) %>% 
  select(-c(Rep1,Rep2,Rep3,Rep4))

BarHCH <- Reps %>% 
  select(Name, Type, Fraction, Compound, Average) %>% 
  pivot_wider(names_from = "Compound", values_from= 'Average') %>% 
  select(Name, Type, Fraction, aHCH, bHCH, gHCH, Sum_HCH) %>%  
  mutate_at(vars(c(aHCH, bHCH, gHCH)), ~(./`Sum_HCH`*100))  %>% 
  mutate(Sum_HCH=aHCH + bHCH +gHCH) %>% 
  select(Name, Type, Fraction, aHCH, bHCH, gHCH) 

BarDDT <- Reps %>% 
  select(Name, Type, Fraction, Compound, Average) %>% 
  pivot_wider(names_from = "Compound", values_from= 'Average') %>% 
  select(Name, Type, Fraction, ppDDE, opDDD, ppDDD, opDDT, ppDDT,Sum_DDTs) %>% 
  mutate_at(vars(c(ppDDE, opDDD, ppDDD, opDDT, ppDDT)), ~(./`Sum_DDTs`*100))  %>% 
  mutate(Sum_DDTs=ppDDE + opDDD + ppDDD + opDDT +ppDDT) %>% 
  select(Name, Type, Fraction, ppDDE, opDDD, ppDDD, opDDT, ppDDT) 

BarChlordane <- Reps %>% 
  select(Name, Type, Fraction, Compound, Average) %>% 
  pivot_wider(names_from = "Compound", values_from= 'Average') %>% 
  select(Name, Type, Fraction, trans_Chlordane, cis_Chlordane, trans_Nonachlor, cis_Nonachlor, Sum_Chlordanes) %>%
  mutate_at(vars(c(trans_Chlordane, cis_Chlordane, trans_Nonachlor, cis_Nonachlor)), ~(./`Sum_Chlordanes`*100))  %>%
  mutate(Sum_Chlordanes=trans_Chlordane + cis_Chlordane + trans_Nonachlor + cis_Nonachlor)  %>% 
  select(Name, Type, Fraction, trans_Chlordane, cis_Chlordane, trans_Nonachlor, cis_Nonachlor)
    
BarPCB7 <- Reps %>% 
  select(Name, Type, Fraction, Compound, Average) %>% 
  pivot_wider(names_from = "Compound", values_from= 'Average') %>% 
  select(Name, Type, Fraction, PCB_28, PCB_52, PCB_101, PCB_138, PCB_153, PCB_180, Sum_PCB7) %>% 
  mutate_at(vars(c(PCB_28, PCB_52, PCB_101, PCB_138, PCB_153, PCB_180)), ~(./`Sum_PCB7`*100))  %>% 
  mutate(Sum_PCB7=PCB_28 +PCB_52+PCB_101+PCB_138+PCB_153+PCB_180) %>% 
  select(Name, Type, Fraction, PCB_28, PCB_52, PCB_101, PCB_138, PCB_153, PCB_180)

BarTotal <- Reps %>% 
  select(Name, Type, Fraction, Compound, Average) %>% 
  pivot_wider(names_from = "Compound", values_from= 'Average') %>% 
  select(Name, Type, Fraction, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_Chlorobenzenes, Sum_PCB, Sum_OC) %>% 
  mutate_at(vars(c(Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_Chlorobenzenes, Sum_PCB)), ~(./`Sum_OC`*100)) %>% 
  mutate(Sum_OC=Sum_HCH + Sum_DDTs + Sum_Chlordanes + Sum_Chlorobenzenes +Sum_PCB) %>% 
  select(Name, Type, Fraction, Sum_HCH, Sum_DDTs, Sum_Chlordanes, Sum_Chlorobenzenes, Sum_PCB) 
    
Barplot5 <- full_join(BarHCH, BarDDT)
Barplot4 <- full_join(BarChlordane, BarPCB7)
Barplot3 <- full_join(Barplot4, Barplot5)
Barplot2 <- full_join(Barplot3, BarTotal)

Barplot <- Barplot2 %>% 
 pivot_longer(cols=-c(Name, Type, Fraction), names_to = "Compound", values_to= 'Percent') %>%
 mutate(Group = case_when(grepl("PCB", Compound) ~ "PCB", grepl("DD", Compound) ~ "DDT", grepl("HCH", Compound) ~ "HCH", grepl("hl",  Compound) ~ "Chlordanes", grepl("Sum", Compound) ~ "Sums"))  %>%
 mutate(Group = ifelse(grepl("Sum",Compound), 'Sums', Group))

Barplot <- full_join(Barplot, Reps2) 
Barplot <- Barplot %>%
 rename(Rep1='Rep1_c', Rep2= 'Rep2_c', Rep3='Rep3_c', Rep4='Rep4_c', ) %>%
 mutate(Type2 = paste0(Type, Fraction)) 
```


```{r Figure 4}

# Stacked barplots showing composition of congeners in each contaminant group

BarSums <-subset(Barplot, Group=='Sums')
BarPest <-subset(Barplot, Group=='Chlordanes')
BarDDT <-subset(Barplot, Group=='DDT')
BarPCB7 <-subset(Barplot, Group=='PCB')
BarHCH <-subset(Barplot, Group=='HCH')
BarPCBs <-subset(Barplot, Group=='Sum_PCBs')

#Sums
BarSums$Name <- factor(BarSums$Name,levels = c("KV", "IF", "GA", "BE","KH"))
BarSums$Type2 <- factor(BarSums$Type2,levels = c("RiverDissolved", "RiverParticulate", "FjordDissolved", "FjordParticulate"))
BarSums$Compound <- factor(BarSums$Compound,levels = c("Sum_PCB", "Sum_DDTs", "Sum_HCH", "Sum_Chlordanes","Sum_Chlorobenzenes"))

ylim.prim_SUMS <- c(0, 100)  
ylim.sec_SUMS <- c(0, 140)    

b <- diff(ylim.prim_SUMS)/diff(ylim.sec_SUMS)
a <- b*(ylim.prim_SUMS[1] - ylim.sec_SUMS[1])

SUMS <- ggplot(BarSums, aes(Name, y = Percent, fill= Compound)) + 
  geom_bar(stat="identity") + 
  scale_fill_manual(values = c("lightpink4","peachpuff3","#B4B7B9","#8B9DAF","#394165","grey13")) + 
  theme_classic() + facet_grid(~Type2) + xlab(NULL) + 
  geom_point(aes(y = a+Rep1*b),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = a+Rep2*b),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = a+Rep3*b),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = a+Rep4*b),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  scale_y_continuous("% OC", sec.axis = sec_axis(~ (. - a)/b, name = "[OC] (pg/L)")) + 
  theme(text = element_text(size=18)) + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
        legend.text=element_text(size=rel(0.8)), legend.key.size = unit(0.25, "cm"))  

#PCB7
BarPCB7$Name <- factor(BarPCB7$Name,levels = c("KV", "IF", "GA", "BE","KH"))
BarPCB7$Type2 <- factor(BarPCB7$Type2,levels = c("RiverDissolved", "RiverParticulate", "FjordDissolved", "FjordParticulate"))
BarPCB7$Compound <- factor(BarPCB7$Compound,levels = c("PCB_28","PCB_52","PCB_101","PCB_118","PCB_138","PCB_153","PCB_180"))

ylim.prim_PCB <- c(0, 100)  
ylim.sec_PCB <- c(0, 30)    

d <- diff(ylim.prim_PCB)/diff(ylim.sec_PCB)
c <- d*(ylim.prim_PCB[1] - ylim.sec_PCB[1])

PCB7 <-ggplot(BarPCB7, aes(Name, y = Percent, fill= Compound)) + 
  geom_bar(stat="identity") +
  scale_fill_manual(values= c("lightpink4","peachpuff3","#B4B7B9","#8B9DAF","#394165","grey13")) + 
  theme_classic() + geom_col() +
  facet_grid(~Type2) + 
  geom_point(aes(y = c+Rep1*d),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = c+Rep2*d),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = c+Rep3*d),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = c+Rep4*d),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  scale_y_continuous("% PCB7", sec.axis = sec_axis(~ (. - c)/d, name = "[PCB7] (pg/L)")) + 
  theme(text = element_text(size=18))+xlab(NULL) + 
  theme(strip.background = element_blank(), strip.text.x = element_blank()) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
        legend.text=element_text(size=rel(0.8)), legend.title = element_blank(), legend.key.size = unit(0.25, "cm"))

#DDT
BarDDT$Name <- factor(BarDDT$Name,levels = c("KV", "IF", "GA", "BE","KH"))
BarDDT$Type2 <- factor(BarDDT$Type2,levels = c("RiverDissolved", "RiverParticulate", "FjordDissolved", "FjordParticulate"))

ylim.prim_DDT <- c(0, 100)  
ylim.sec_DDT <- c(0, 40)   

f <- diff(ylim.prim_DDT)/diff(ylim.sec_DDT)
e <- f*(ylim.prim_DDT[1] - ylim.sec_DDT[1])

DDT <-ggplot(BarDDT, aes(Name, y = Percent, fill= Compound)) + 
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("lightpink4","peachpuff3","#B4B7B9","#8B9DAF","#394165","grey13")) + 
  theme_classic() + geom_col() + facet_grid(~Type2) + 
  geom_point(aes(y = e+Rep1*f),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = e+Rep2*f),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = e+Rep3*f),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  geom_point(aes(y = e+Rep4*f),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  scale_y_continuous("% DDTs", sec.axis = sec_axis(~ (. - e)/f, name = "[DDTs] (pg/L)")) + 
  theme(text = element_text(size=18)) + xlab(NULL) + 
  theme(strip.background = element_blank(), strip.text.x = element_blank()) + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        legend.text=element_text(size=rel(0.8)), legend.key.size = unit(0.25, "cm"), legend.title = element_blank())

#HCH
BarHCH$Name <- factor(BarHCH$Name,levels = c("KV", "IF", "GA", "BE","KH"))
BarHCH$Type2 <- factor(BarHCH$Type2,levels = c("RiverDissolved", "RiverParticulate", "FjordDissolved", "FjordParticulate"))

ylim.prim_HCH <- c(0, 100)  
ylim.sec_HCH <- c(0, 100)    

h <- diff(ylim.prim_HCH)/diff(ylim.sec_HCH)
g <- h*(ylim.prim_HCH[1] - ylim.sec_HCH[1])

HCH <- ggplot(BarHCH, aes(Name, y = Percent, fill= Compound)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values= c("lightpink4","peachpuff3","#8B9DAF","#394165")) +
  theme_classic() + geom_col() + theme_classic() + xlab(NULL) +
  facet_grid(~Type2) +
  geom_point(aes(y = g+Rep1*h),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  geom_point(aes(y = g+Rep2*h),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  geom_point(aes(y = a+Rep3*b),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  geom_point(aes(y = g+Rep4*h),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  scale_y_continuous("% HCHs", sec.axis = sec_axis(~ (. - g)/h, name = "[HCHs] (pg/L)")) + 
  theme(text = element_text(size=18)) + 
  theme(strip.background = element_blank(), strip.text.x = element_blank()) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        legend.text=element_text(size=rel(0.8)), legend.key.size = unit(0.25, "cm"), 
        legend.title = element_blank())

#Pesticides
BarPest$Name <- factor(BarPest$Name,levels = c("KV", "IF", "GA", "BE","KH"))
BarPest$Type2 <- factor(BarPest$Type2,levels = c("RiverDissolved", "RiverParticulate", "FjordDissolved", "FjordParticulate"))

ylim.prim_Pest <- c(0, 100)  
ylim.sec_Pest <- c(0, 10)    #10 

j <- diff(ylim.prim_Pest)/diff(ylim.sec_Pest)
i <- j*(ylim.prim_Pest[1] - ylim.sec_Pest[1])

Pest <- ggplot(BarPest, aes(Name, y = Percent, fill= Compound)) + 
  geom_bar(stat="identity")+scale_fill_manual(values =c("lightpink4","peachpuff3","#8B9DAF","#394165")) +
  theme_classic() + 
  facet_grid(~Type2) + geom_col() + xlab(NULL) +
  geom_point(aes(y = i+Rep1*j), shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = i+Rep2*j),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = i+Rep3*j),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) + 
  geom_point(aes(y = i+Rep4*j),shape = 21,colour = "white", fill = "black", size = 4, stroke = 0.5) +
  scale_y_continuous("% Chlordanes", sec.axis = sec_axis(~ (. - i)/j, name = "[Chlor.] (pg/L)")) + 
  theme(text = element_text(size=18)) + 
  theme(strip.background = element_blank(),strip.text.x = element_blank(), 
        legend.text=element_text(size=rel(0.8)), legend.key.size = unit(0.25, "cm"), legend.title = element_blank())


# Compile figures
Figure4 <- SUMS/PCB7/DDT/HCH/Pest

# Save plot
ggsave("Figure4.pdf", Figure4 , width = 20, height = 18, device = cairo_pdf)
```


## Figure 5. Isomer Ratios

```{r Figure 5}

# Pesticide Isomer ratios. 

Isomer_Ratios <- Data %>% 
  filter(Volume_L != 1) %>% 
  mutate(Ratio_HCH=(aHCH*100)/(gHCH*100)) %>% 
  mutate(Ratio_Chlor=(trans_Chlordane*100)/(cis_Chlordane*100)) %>% 
  mutate(Ratio_Nona=(trans_Nonachlor*100)/(cis_Nonachlor*100)) %>% 
  mutate(Ratio_DDT=((opDDD+ppDDD+ppDDE)*100)/((opDDT+ppDDT)*100)) %>% 
  select(Station, Type, Fraction, Ratio_HCH, Ratio_Chlor, Ratio_Nona, Ratio_DDT) 

Isomer_Ratios$Type <- factor(Isomer_Ratios$Type,levels = c("River", "Fjord"))
Isomer_Ratios$Station <- factor(Isomer_Ratios$Station,levels = c("KV", "IF", "GA", "BE", "KH"))

Plot1 <- Isomer_Ratios %>%
  ggplot(aes(x=Type, y=Ratio_HCH, fill = Fraction)) + 
  geom_boxplot(outliers = FALSE, outlier.shape = NA) + 
  xlab(NULL) + scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + 
  geom_point(position=position_jitterdodge(), stat='identity',shape = 21, size = 2,alpha=1) +
  ylab(expression(paste(alpha,'-HCH: ',gamma,'-HCH'))) + xlab(NULL) + theme(text = element_text(size=14), legend.position = "none")

Plot2 <- Isomer_Ratios %>%
  ggplot(aes(x=Type, y=Ratio_Chlor,fill=Fraction)) + 
  geom_boxplot(outliers = FALSE, outlier.shape = NA) + 
  xlab(NULL) + scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + 
  theme_classic() + 
  geom_point(position=position_jitterdodge(), stat='identity',shape = 21, size = 1.5,alpha=1) + 
  ylab(expression(italic("trans:cis-")-"chlordane")) + xlab(NULL) + theme(text = element_text(size=14), legend.position = "none") + ylim(0,1.5)

Plot3<- Isomer_Ratios %>%
  ggplot(aes(x=Type, y=Ratio_Nona,fill=Fraction)) + 
  geom_boxplot(outliers = FALSE, outlier.shape = NA) + xlab(NULL) +
  scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + theme_classic() + 
  theme(legend.position='none') +
  geom_point(position=position_jitterdodge(), stat='identity',shape = 21, size = 2,alpha=1) + 
  ylim(0.5,2) + ylab("trans:cis Nona") + xlab(NULL) + theme(text = element_text(size=18))

Plot4 <- Isomer_Ratios %>%
  ggplot(aes(x=Type, y=Ratio_DDT,fill=Fraction)) + 
  geom_boxplot(outliers = FALSE, outlier.shape = NA) + xlab(NULL) +
  scale_fill_manual(values=c("#8B9DAF","lightsalmon4")) + theme_classic() + 
  geom_point(position=position_jitterdodge(), stat='identity',shape = 21, size = 2,alpha=1) + 
  ylab("(DDD+DDE)/DDT") + xlab(NULL) + theme(text = element_text(size=14))

# Print Figure
p <- (((Plot1|Plot2)|Plot4) + plot_annotation(tag_levels = 'A'))

ggsave("Figure5.pdf", p, width = 10, height = 3.5, device = cairo_pdf)
```

``` {r Ratio Tests}

# Wilcoxon rank sum tests of isomer ratios between river and fjord stations for each fraction separately. 

Dis <- Isomer_Ratios %>%
  filter(Fraction=='Dissolved')  %>%
  select(-Fraction) %>% 
  pivot_longer(cols=-c(Type, Station), names_to = "Compound", values_to= 'Concentration') %>%   
  nest(-Compound) %>%                  
  mutate(results = map(data, ~tidy(wilcox.test(Concentration ~ Type, data = .)))) %>%
  select(-c(data)) %>%
  unnest(results)

Part <- Isomer_Ratios %>%
  filter(Fraction=='Particulate')  %>%
  select(-Fraction) %>% 
  pivot_longer(cols=-c(Type, Station), names_to = "Compound", values_to= 'Concentration') %>%   
  nest(-Compound) %>%                  
  mutate(results = map(data, ~tidy(wilcox.test(Concentration ~ Type, data = .)))) %>%
  select(-c(data)) %>%
  unnest(results)

Dis %>% 
  select(-c(method, alternative))  %>% 
  kable(caption= 'Dissolved: Results of Wilcoxon Rank Sum Two-sided Test btw Fjord and River Dissolved samples', digits=4) %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE)

Part  %>% 
  select(-c(method, alternative))  %>% 
  kable(caption= 'Particulate: Results of Wilcoxon Rank Sum Two-sided Test btw Fjord and River Particulate samples', digits=4) %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = TRUE)

# Table with summary stats
Table <- Isomer_Ratios %>% 
  pivot_longer(cols= -c(Station, Type, Fraction), names_to = "Compound", values_to= 'Concentration') %>% 
  group_by(Compound, Type, Fraction)  %>% 
  arrange(Compound, Type, Fraction)  %>% 
  summarize(n=n(), Mean=mean(Concentration, na.rm=T), stdev= sd(Concentration, na.rm=T), min = min(Concentration), max = max(Concentration)) 

Table %>% 
  kable(caption= 'Pesticide Isomer Ratios', digits=2, booktabs = T, align = c("r"))  %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```


## Figure 6. Ratio (Part: Diss) vs. Log KOW for each compound. 

```{r Figure 6}

# Plot partitioning (part:Diss) vs LogKOW of each contaminant

Ratio2 <- Data %>%
  select(-c(PCB_18, PCB_31, PCB_33, PCB_47, PCB_66, PCB_74, PCB_99, PCB_105, 
            PCB_118, PCB_149, PCB_187, Latitude, Longitude, Name)) %>%
  arrange(Lab_ID, Volume_L) %>%
  ungroup() %>%
  pivot_longer(cols=-c(Lab_ID, Station, Type, Fraction, Volume_L), 
               names_to = "Compound", values_to= 'Concentration') %>%
  mutate(Reps= 'Rep1') %>%
  mutate(Reps= ifelse(grepl("3075|3076|3083|3084|3091|3092|3095|3096|3099|3100|3107|3108|3113|3114",Lab_ID),
                      'Rep2', Reps)) %>%
  mutate(Reps= ifelse(grepl("3077|3078|3085|3086|3109|3110",Lab_ID), 'Rep3', Reps)) %>%
  mutate(Reps= ifelse(grepl("3079|3080|3087|3088",Lab_ID), 'Rep4', Reps)) %>%
  select(-c(Lab_ID, Volume_L)) %>%
  filter(!Compound %in% c("Heptachlor", "Mirex","Sum_PCB", "Sum_PCB7","Sum_Chlordanes",
                          "Sum_HCH", "Sum_DDTs", "Sum_Chlorobenzenes", "Sum_OC")) %>%
  pivot_wider(names_from="Fraction", values_from="Concentration") %>%
  mutate(Ratio=(Particulate*100)/(Dissolved*100)) %>%
  ungroup()
  
Ratio <- full_join(Ratio2, LogKow) %>% drop_na(Station) %>% 
  group_by(logKOW)

Ratio$logKOW <- as.factor(Ratio$logKOW)
Ratio$Type <- factor(Ratio$Type,levels = c("Fjord", "River"))

Ratio_plot <- ggplot(Ratio, aes(x=logKOW, y=Ratio, fill=Type)) + 
  geom_boxplot(outlier.shape = NA) + 
  xlab("Log Kow") + 
  ylab("Particulate:Dissolved") + 
  theme_classic() + 
  scale_fill_manual(values= c("#8B9DAF","peachpuff3")) + 
  geom_point(position=position_jitterdodge(), stat='identity',shape = 21, size = 2,alpha=1) + 
  scale_y_log10() + ylim(0,4) + theme(text = element_text(size=17)) +
  geom_hline(yintercept=1,linetype="dashed")

# Print Figure
ggsave("Figure6.pdf", Ratio_plot, width = 11, height = 4, device = cairo_pdf)
```


## Figure S2. TS diagram

```{r Figure S2}
# TS diagram in Mikko's PlotSvalbard package

CTD$Name <- factor(CTD$Name,levels = c("IF", "GA", "BE","KH"))

FigureS2 <- ts_plot(CTD, temp_col = "Temp", sal_col = "Salinity",
                    WM = kongsfjord_watermasses, color_wmpoly = "black", color = "Name", zoom = TRUE, nlevels = 6,
                    color_scale = c("cyan3","burlywood4","darkolivegreen","deepskyblue4"),
                    symbol_shape=16, symbol_size = 4,
                    symbol_alpha = 1, color_var_name = NULL,
                    margin_width = 0.15, margin_height = 0.2,
                    plot_data = TRUE, base_size = 12)

# Save plot
ggsave("FigureS2.pdf", FigureS2, width = 5, height = 5, device = cairo_pdf)
```

